---
title: "assesmentOfGeneSets"
author: "Viktor Hesselberg-Thomsen"
date: "29/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(data.table)
library(KEGGREST)
library(ggplot2)
library(plotly)
library(htmlwidgets)
source('custom_functions.R', local = knitr::knit_global())
```

```{r}
# set bacteria
bac <- 'P'
```


```{r}
# load data
geneSets_kobas <- readRDS(file = here(paste0('data/tidy/geneSets/geneSets_Kobas_', bac, '.rds')))
geneSets_eggnog <- readRDS(file = here(paste0('data/tidy/geneSets/geneSets_', bac, '.rds')))
richfactor_kobas <- read.csv(file = here(paste0('data/tidy/geneSetEnrichmentAnalysis/richFactor_kobas_', bac, '.csv')))
richfactor_eggnog <- read.csv(file = here(paste0('data/tidy/geneSetEnrichmentAnalysis/richFactor_eggnog_', bac, '.csv')))

```

Investigate some summary stats of the gene sets sizes
```{r}
# number of gene set assocciations 
print('eggnog')
sum(richfactor_eggnog$size)
summary(richfactor_eggnog$size)
print('number of gene sets')
length(richfactor_eggnog$size)

print('kobas')
sum(richfactor_kobas$size)
summary(richfactor_kobas$size)
print('number of gene sets')
length(richfactor_kobas$size)

# boxplot
boxplot(data.frame('eggnog' = richfactor_eggnog$size, 'kobas' = richfactor_kobas$size), outline = F)

```

* There are way fewer genes associated to gene sets in kobas (for P). Thus, maybe the each gene is associated with fewer gene sets.
* The small median of eggnog could indicate many incomplete or falsely assigned gene sets. This is supported by the large number of gene sets in eggnog based annotation
* Actually >25 of the gene sets from eggnog will be removed with the size > 5 threshold.
* Based on only the above results, so far kobas seem the most reliable

```{r}
geneOccurences <- merge(x = as.data.frame(table(unlist(geneSets_eggnog))),
                        y = as.data.frame(table(unlist(geneSets_kobas))),
                        by = 'Var1', all = T, suffixes = c('_eggnog', '_kobas'))
colnames(geneOccurences)[1] <- 'geneID'
dim(geneOccurences)
head(geneOccurences)

ggplot(data = geneOccurences)+
  geom_point(aes(Freq_eggnog, Freq_kobas)) 

```




